Technology-based aids for people affected by Autism Spectrum Disorder (ASD)

### 2017-11-17 paper.tex
Concluded first part in diagnostic

%% Glossary
\newacronym{asd}{ASD}{Autism Spectrum Disorder}
\newacronym{va}{VA}{Virtual Agents}
\newacronym{dof}{DoF}{Degrees of Freedom}

@@

% TODO Brief presentation of ASD (?)
\gls{asd} is a term used to cover a very big set of disorders. In this field there are a lot of studies and experiments, but still is one of the most unknown \DP{forse unknown non è il massimo come termine qui} disease. The symptoms are well known, while the causes are mostly unknown. Nowadays, \textit{screening tests} are used to classify a person as affected by \gls{asd}, but these tests have an high percentage of error (false positive or negative) and they can be administered on a patient at least 3 years old. For this reason, several techniques have been studied to wonder if a children has this disorder in the first 2 years of life.

In this respect, this work is divided in two main sections. In \autoref{sec:thera_tech} we present some innovative tool to improve social skills of a child affected by ASD faster than a classical face-to-face therapy will does \DP{Questa frase fa schifo}.
In \autoref{sec:every_tech} we present a couple of wearable products that allow a faster and easier integration of the subject in the everydays tasks.

The classical way to conduct a therapy session, is a face-to-face meeting between the subject and the therapist.
Usually, it is conducted in four phases: the first one is called \textit{Instruction}, in which the therapist explains what is the skill that the subject is going to learn; the second one is the \textit{Modeling} phase, in which the therapist shows the patient what he has to do; there is then the \textit{Rehearsal} phase, in which the patient tries to imitate the therapist and finally a \textit{Feedback} phase in which the therapist draws its own conclusions about the test.
The problem of this kind of therapy are the completely subjective evaluation procedure carried out by the therapist and that the subject's attention is not immediate.

%TODO Avatar-based techniques to improve the performances of the therapy
%TODO Methods to detect the anxiety level
To improve the performances of the sessions, the use of \gls{va} platforms has been introducted.
A \gls{va} platform consist of an avatar, that can be a robot or a virtual character, a learning software that receive information from the Avatar, elaborate them and adjust the response of the Avatar, and an interface for the therapist to interpretate the results.
In a \gls{va} session, the therapist, with the aids of the learning software, learns the Avatar in order to do an action that the child has to imitate, the child tries to imitate the Avatar learning new way to interact with the society.
Different system has been proposed, we propose here two different solutions, one more general and one more complex and efficient.

The idea of an Avatar-based therapy session is proposed firstly in \cite{Guerrero-vasquez}. They propose a system composed by a \gls{va} and an \textit{expert system} used by the therapist to instruct the \gls{va}. A general scheme of the proposed system is shown in \autoref{fig:avatar_scheme}.

They chose to use a Virtual Agent instead of a robot since, according with the needs, it provides the possibility to customize the tool, while a robot is limitated on its shape and its \gls{dof}.

As can be seen in \autoref{fig:avatar_tasks}, the platform is organized as follow:
\begin{LaTeXdescription}
	\item[Child Interface] It selects the child's favourite Virtual Agent and uses interaction tools as screen, camera, microphone and speakers to interact with the child;
	\item[Expert System] Contains a learning algorithm that promotes interaction between the child and the Virtual Agent, it stores the scenario and features of the Virtual Agent and it monitors progress of the session in order to generate statistical reports that will be used by the therapist;
	\item[Therapist Interface] Allow the therapist to access to the reports generated by the Expert System and gives information about the current therapy, in order to allow the therapist to draw conclusions.
\end{LaTeXdescription}

The choices about the best \gls{va} to use during the sessions, or the characteristics that has to be changed in the current \gls{va}, are taken monitoring the reaction of the child when he's put in front of them.
The character can be not strictly a human, since the highest degree of affinity can be towards an animal or an object.

The key point in their work is that the information received on the screen of the therapist interface shows a real time image of the child and monitored progress statistics.
Since the VA that the child observes on his screen is a disguise of his therapist, combined with artificial intelligence, is actually the therapist who promotes the interaction in combination with the expert system.

Another more complex and complete approach was performed in \cite{Alahbabi17}, using a robot instead of a \gls{va}, since it has been shown in \cite{Pioggia08} that patients show a more favorable response towards humanoid robots than to other non-humanoid virtual characters.

This approach has sensible advantages but also some intrinsic shortcomings:
\begin{itemize}
	\item most of the robotic or virtualized therapy are preprogrammed, that make the session boring and repetitive;
	\item robots are limited and work in a controlled environment;
	\item usually robots don't have a quantitative behavioral analysis tool due to the absence of therapists during their development process.
\end{itemize}

@@

\subsection{Hand-eye coordination}
\label{sec:hand_eye}
A different aspect to take into consideration when facing subjects with ASD, is the motor disorder related to the syndrome. In \cite{Casellato2017} a new approach to face the issue is presented, where the proposed protocol may be used for ad hoc subject training, helping children with ASD to enhance motor abilities related to hand-eye coordination.

@@

Figures \ref{task1} and \ref{task1_pT} show the results just described.

@@

\section{Everydays tools techniques}
\label{sec:every_tech}

\subsection{Jumping Ball}
\subsection{Genetic Algorithm}

>>>>

%% Glossary
\newacronym{asd}{ASD}{Autism Spectrum Disorder}
\newacronym{va}{VA}{Virtual Agents}
\newacronym{dof}{DoF}{Degrees of Freedom}
\newacronym{td}{TD}{Typically-Developing}

@@

% TODO Brief presentation of ASD (?)
\gls{asd} is a term used to cover a very big set of disorders. There are a lot of studies and experiments in this field, but still a lot questions about it are unresolved \DP{forse unknown non è il massimo come termine qui} \cri{Prova a vedere se così ti piace}: for example, the symptoms are well known, while the causes are still mostly unknown. Nowadays \textit{screening tests} are used to classify a person as affected by \gls{asd}, but these tests have an high percentage of error (false positive or negative) and they can be administered to a patient who is at least 3 years old. For this reason, several techniques have been studied in order to establish if a children has this disorder in the first 2 years of life.

With respect to these aims, this work is divided into two main sections: In \autoref{sec:diagnosis} we present some innovative tools and procedures to help diagnosing the syndrome in the first one or two years of a child; in \autoref{sec:thera_tech} we describe three \cri{vedere se poi il numero va bene} devices and methods useful for learning and improving cognitive skills.

@@

\subsection{Hand-eye coordination}
\label{sec:hand_eye}
A different aspect to take into consideration when facing subjects with \gls{asd}, is the motor disorder related to the syndrome. In \cite{Casellato2017} a new approach to face the issue is presented, where the proposed protocol may be used for ad hoc subject training, helping children with ASD to enhance motor abilities related to hand-eye coordination.

@@

Figures \ref{fig:task1} and \ref{fig:task1_pT} show the results just described.

@@

During the second task there is a significant difference between \gls{td} (TD) children and children with ASD and between the first and the second attempt:
\begin{itemize}
\item Even though there is not a strong correlation among the children of a same group, it can be noticed that both TD children and ASD children performed better during the second attempt, when they already knew the path to follow a priori;
\item In general, \gls{td} children had better performance than \gls{asd} children;
\item For ASD children, the \textit{hand-gaze} distance was higher than for TD children, but quite the same in both the attempts.
\end{itemize}

Figure \ref{path} shows an example of trial with a TD child.

\begin{figure}
\centering
\includegraphics[width = .55\textwidth]{path.png}
\caption{Trials A and B for a \gls{td} child}
\label{path}
\end{figure}

\subsubsection{summary observations}
\label{sec:summary}

The experiment results can be summarized with a few crucial conclusions:
\begin{itemize}
\item the first task showed how children with \gls{asd} took longer to learn the task execution correctly, even though once the process had been assimilated the execution was correct, this meaning that these children showed a delayed ability to learn from sensory prediction errors; moreover the ASD child had more variability in motor execution, this suggesting more difficulty in maintaining performance consistency, and the gaze pattern was sticking less to the hand pattern if compared to the \gls{td} children.
\item the second task showed how children with ASD obtained performance slightly worse than the TD children, but anyway they were able deploy the information regarding the previously assimilated path to improve the performance. This suggests that gls{asd} children are able to take advantage of the availability of the whole motor plan and supports the hypothesis that simple motor planning is intact but there is a diminished use of feedback.
\end{itemize}

From these results, since the timing (which is strictly associated with adaptation and prediction) is one of the crucial aspects managed by the cerebellum (known to be the cerebral section reporting the most consistent brain differences in autism), it can be supported the hypothesis that bad timing in sensory integration causes degraded motor performance and delayed learning in children with \gls{asd}.

Such dysfunctions could affect not only motor processes but also cognitive ones, and this form of motor tracking could also be useful to better understand the mechanism at the basis of the well-documented visuo-motor impairments in ASD (\cite{Chow2010}, \cite{Green2013}, \cite{Papa}) and throw more light on the functional integrity of brain networks during development in people with ASD.

The classical way to conduct a therapy session is a face-to-face meeting between the subject and the therapist.
Usually, it is conducted in four phases: the first one is called \textit{Instruction}, in which the therapist explains what is the skill that the subject is going to learn; the second one is the \textit{Modeling} phase, in which the therapist shows the patient what he has to do; then there is the \textit{Rehearsal} phase, in which the patient tries to imitate the therapist, and finally a \textit{Feedback} phase in which the therapist draws its own conclusions about the test.
The problem of this kind of therapy structure is the completely subjective evaluation procedure carried out by the therapist and the non-immediate subject's attention.

%TODO Avatar-based techniques to improve the performances of the therapy
%TODO Methods to detect the anxiety level
To improve the performance of the sessions, the use of \gls{va} platforms has been introducted.
A \gls{va} platform consists of an avatar, that can be a robot or a virtual character, a learning software that receives information from the Avatar, elaborates them and adjusts the response of the Avatar, and an interface for the therapist that interprets the results.
In a \gls{va} session the therapist, with the aids of the learning software, trains the Avatar to do an action that the child has to imitate. At the same time, the child tries to imitate the Avatar learning new ways to interact with the society.
Different system has been proposed, we describe here two different solutions, one more general and one more complex and efficient.

The idea of an Avatar-based therapy session is shown firstly in \cite{Guerrero-vasquez}. In the paper they propose a system composed by a \gls{va} and an \textit{expert system} used by the therapist to instruct the \gls{va}. A general scheme of the system is shown in \autoref{fig:avatar_scheme}.

They chose to use a Virtual Agent instead of a robot since, according with the needs, it provides the possibility to customize the tool, while a robot is limited by its structure and its \gls{dof}.

As can be seen in \autoref{fig:avatar_tasks}, the platform is organized as follows:
\begin{LaTeXdescription}
	\item[Child Interface] It selects the child's favorite Virtual Agent and uses interaction tools as screen, camera, microphone and speakers to interact with the child;
	\item[Expert System] Contains a learning algorithm that promotes interaction between the child and the Virtual Agent, it stores the scenario and features of the Virtual Agent and it monitors progress of the session in order to generate statistical reports that will be used by the therapist;
	\item[Therapist Interface] Allow the therapist to access to the reports generated by the Expert System and gives information about the current therapy, in order to allow the therapist to draw conclusions.
\end{LaTeXdescription}

The choices about the best \gls{va} to use during the sessions, or the characteristics that has to be changed in the current \gls{va}, are taken monitoring the reaction of the child when he is put in front of them.
It is not necessary for the character to be a human, since the highest degree of affinity can be towards an animal or an object.

The key point in this work is that the information received on the screen of the therapist interface shows a real time image of the child and monitored progress statistics.
Since the VA that the child observes on his screen is his therapist in disguise, actually it is the therapist who, combined with artificial intelligence, promotes the interaction in combination with the expert system.

Another more complex and complete approach was performed in \cite{Alahbabi17}, using a robot instead of a \gls{va}; this because it has been shown in \cite{Pioggia08} that patients manifest a more favorable response towards humanoid robots than to other non-humanoid virtual characters.

This approach has sensible advantages but also some intrinsic shortcomings:
\begin{itemize}
	\item most of the robotic or virtualized therapies are preprogrammed and this makes the session boring and repetitive;
	\item robots are limited and work in a controlled environment;
	\item usually robots do not have a quantitative behavioral analysis tool, due to the absence of therapists during their development process.
\end{itemize}
